\documentclass[12pt,a4paper,twoside]{report}
\usepackage{amsmath}
\usepackage{bm}
\usepackage[colorlinks=true,urlcolor=blue]{hyperref}
\usepackage[table]{colortbl}
\usepackage{graphicx}
\usepackage{fullpage}
\title{Snp HeRitability Estimation Kit (SHREK) Manual}
\date{\today}
\author{Sam Shing Wan, Choi\\ 
\texttt{choishingwan@gmail.com}
\and
Johnny Sheung Him, Kwan\\
\texttt{shkwan@hku.hk}
\and
Henry Chi-Ming, Leung\\
\texttt{cmleung2@cs.hku.hk}
}

\begin{document}

\maketitle
\tableofcontents
\chapter{To Do List}
\section{Programme}
If time allows, re-structure some of the scripts.
\begin{enumerate}
\item{Write a better make file. Maybe also include the configure, if I can figure out how to write those}
\end{enumerate}
\section{Documentation}
Finish this read me document
\chapter{Background}
\chapter{Installation}
\section{Dependencies}
The programme is dependent on the ISO C++ 2011 standard and Eigen library.
\section{Tutorial}
\chapter{Input formats}
\section{Test statistic file}
\section{Linkage file}
The linkage file should be in the binary ped format. 
\chapter{Examples}
\chapter{Assumptions and limitations}
\section{Assumptions}
\subsection{Calculation of variance}
There are multiple assumptions when we calculate the variance. 
First, we assume a big sample size. Only when the sample size is big, can we approximate the distribution of the effects to follow the $\chi^2$ distribution. 
Therefore, when the sample size is small, our variance estimate can be biased.

Secondly, for simplicity of implementation, when calculating the variance covariance matrix of the effects, we uses the \textbf{maximum sample size} as the $n$.
So, for example, if 9 out of 10 Snps' association were conducted on 100 samples, yet the remaining 1 Snp were conducted on 10,000 samples, the programme will still use $n= 10000$ in the calculation of the variance covariance of \textbf{all} the Snp pairs.
\section{Limitations}
\subsection{Direction of effect}
If one doesn't provide the direction of effect, the programme will use the direction from the test-statistic. 
However, if only p-values are provided, all effect will be assumed to be positive. 
This will lead to a positive bias in the variance explained.
Therefore it is advised to always provide the direction of effect.
\subsection{Perfect LD}
In the case where two or more Snps are in perfect LD, we cannot determine the true underlying heritability explained of each Snps.
Therefore, in SHREK, Snps in perfect LD will be grouped together and formed one Snp where its effect is the mean of all Snps in the group.
Decomposition will then be performed on this Snp and the resulting heritability will be \textbf{equally distributed} among the original group of Snps.

So for example:
Assuming there are 5 Snps, $Snp_1, Snp_2, Snp_3, Snp_4 and Snp_5$ where $Snp_1$, $Snp_2$ and $Snp_3$ are all in perfect LD each with effect $f_1$,$f_2$ and $f_3$ respectively.
Then we will first get $Snp_\mu$ with effect as $f_\mu = \frac{f_1+f_2+f_3}{3}$. 
As $Snp_1$, $Snp_2$ and $Snp_3$ has the same LD structure with all other Snps (as they are in perfect LD), we then use the LD structure of $Snp_1$ as a representative of the LD of $Snp_\mu$ with all other Snps. 
This will then gives us
\[ \left( \begin{array}{ccc}
R_{11} & R_{14} & R_{15}\\
R_{14} & R_{44} & R_{45}\\
R_{15} & R_{45} & R_{55}
\end{array} \right)
%
\left( \begin{array}{c}
h_\mu \\
h_d \\
h_e
\end{array} \right)
=
\left( \begin{array}{c}
f_\mu \\
f_d \\
f_e
\end{array} \right)
\]
Finally, we will get
\begin{align*}
h_1 &= \frac{h_\mu}{3} \\
h_2 &= \frac{h_\mu}{3} \\
h_3 &= \frac{h_\mu}{3}
\end{align*}

It is important to note that for the variance, it is \textbf{NOT} equally divided among the Snps. 
Instead, it is something like

\begin{align*}
var(h_1) &= var(h_\mu) \\
var(h_2) &= var(h_\mu) \\
var(h_3) &= var(h_\mu)
\end{align*}
\subsection{Per Snp Heritability}
Although the programme does produce a per snp heritability estimation, this value only served as an approximation.
For example, if two causal snps are in perfect LD and each has a different effect size, then per snp heritability will not work as the effect will be equally divided among them.
We do expect similar situation will occur when there is LD between the causal snps.
The total heritability estimation will still work, but not the per snp effect size estimation.
Therefore users should be cautious when they try to interpret the per snp effect.
\end{document}